---
title: "Getting Down with R Markdown"
author: "Kaylene McClanahan"
date: "December 13, 2017"
output: 
  ioslides_presentation:
    widescreen: true
    transition: faster
---

## Hi, I'm Kaylene

### Professional background:
* PhD student in Management and Organizations at the Kellogg School of Management 
* Research in hierarchy and leadership emergence
* Have been teaching myself R for about a year and a half
* Going into data science or analytics post graduation

### Other interests:
1. Being outside
2. Eating eggs
3. Finger guns
4. Babies  





## What is R Markdown?
* A way of taking your R code and/or results and getting them somewhere else
* Great for reporting results, putting them on the web, etc.
* Also a great to way to organize and annotate projects

## What can you do with R Markdown?
* Make presentations (like this one!)
* Make pdfs, word documents, or basically any other kind of document
* Make your work more reproducible
* Make interactive (or not!) webpages
* Write books  


## Are there any drawbacks to R Markdown?
* Formatting is more limited
* Can be more time consuming up front
* There is a little bit of a learning curve  

## How to Use R Markdown 
1. Install 'rmarkdown' package---should automatically install other dependent packages 
```{r, eval = FALSE}
install.packages('rmarkdown')
```
2. In Rstudio, use drop-down menu to create a new .Rmd file 
3. Make your report/document/webpage (more details soon) 
4. "Knit" your document into a new format 

## Every R Markdown document has two components
1. **Header:** this tells the R Markdown document telling it what it's going to be become 
2. **Body:** This is where you type what are going to publish, usually includes text and code from R 


## About those headers
* Written in YAML (Yet Another Markup Language)
* Will auto-populate if you use the dropdown menu in RStudio

```{r, eval = FALSE}
---
title: "Getting Down with R Markdown"  
author: "Kaylene McClanahan"  
date: "December 13, 2017"  
output: ioslides_presentation
---
```


## About those headers, con't
* Can use to make document-wide style choices
```{r, eval = FALSE}
---
title: "Getting Down with R Markdown"  
author: "Kaylene McClanahan"  
date: "December 13, 2017"  
output:  
  ioslides_presentation:  
  widescreen: true  
  transition: faster  
---
```


## About that body
* You will very likely include narration; this is written mostly normally 
* Formatting choices are going to be written into the text (but this doesn't interfere with readability) 

\#\# Every R Markdown document has two components  
\* \*\*Header:\*\* this tells the R Markdown document telling it what it's going to be become  
\* \*\*Body:\*\* This is where you type what are going to publish, usually includes text and code from R  

## Every R Markdown document has two components
* **Header:** this tells the R Markdown document telling it what it's going to be become  
* **Body:** This is where you type what are going to publish, usually includes text and code from R 

##Some R Markdown Formatting Basics
The very essentials:

\#\# This is a header  
\#\#\# This is a smaller header  
Use single backticks for package names, etc.: `dplyr`  
This is how you do \*\*italics\*\* and \*bold\*   
End a line with two spaces to start a new paragraph.    

becomes:  

## This is a header  
### This is a smaller header   
Use single backticks for package names, etc.: `dplyr`  
This is how you do *italics* and **bold**.   
End a line with two spaces to start a new paragraph.    


## But what about...R?
* Embed chunks code from R using three backticks and curly bracket  

\`\`\`\{r\}  
x <- 3  
y <- 3 + 4  
\`\`\`  
becomes: 
```{r}
x <- 3
y <- 3 + 4
```
## Snippets
* Embed snippets into your text using the a single backtick and a lowercase r.  

There were a total of \`\r r 3 + 5\` paragraphs in the report.

becomes:

There were a total of `r 3 + 5` paragraphs in the report.

## More advanced embedding options:
Do you want to include the results, but not code?  
\`\`\`\{r, echo = FALSE\}

Do you want to show the code but not run it?  
\`\`\`\{r, eval = FALSE\}

Do you want messages to not appear?  
\`\`\`\{r, message = FALSE\}

## And then, you knit!
* In R Studio, this is as easy as clicking a button  
* In R, you run:  
 ```{r, eval = FALSE}
 rmarkdown::render("<file path>") 
 ```

## You are approached by a NYC start-up: "Chugging Along"
* Hangover-preventation-focused Uber competitor (make passengers chug water before exiting the car)
* Want to know everything they can about parties in NYC
* They give you some (seemingly v flawed) data
* V strapped for cash---can only afford to hire you for ~ 45 mins of work
* Want to see your process, but in a digestable output (R Markdown???)

## Party Data
* Entirety of the party-related noise-complaints in 2016 in NYC  
    * From [Kaggle](https://www.kaggle.com/somesnm/partynyc)
    * Also includes time/date of call, time/date it was closed, borough, zip code,        location type  
* Joined this with some NYC demographic data from the 2010 census  
    * From [Baruch College Library](https://www.baruch.cuny.edu/confluence/display/geoportal/NYC+Geographies) (thanks, Frank Donnelly!)
    * Demographic information is at the PUMA level  
    * PUMAs are public use microdata areas. They are roughly 100K people.  
    * Includes a variety of things that I thought were interesting  

## Part(l)y Data?
* There are some issues 
* There is the party_ready data frame, or the party_sideways which is limited to 55   unique observations

## Now go explore and report! 
* Each group has an R Markdown cheat sheet and list of variables
* Spend a few minutes talking about what you'd like to do
* Spend about 20 minutes playing around with the data
* Spend about 20 minutes putting your results in R Markdown


